---
title: "Global Plastic Pollution"
description: |
  Tidy Tuesday Global Plastic Pollution Data Visualization
author:
  - name: Danielle  Sclafani
    url: {}
date: 03-07-2021
output:
  distill::distill_article:
    self_contained: false
    code_folding: hide
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(patchwork)
library(ggalt)
library(sf)
library(ggtext)
library(here)
library(tmap)
library(leaflet)
library(tigris)
```

## Tidy Tuesday Data Visualization

Using the data provided by Tidy Tuesday, change in global plastic pollution from 2019-2020 is explored for the ten countries with the greatest amount of plastic pollution.


### Reading in Data from TidyTuesday
```{r}
tuesdata <- tidytuesdayR::tt_load('2021-01-26')
tuesdata <- tidytuesdayR::tt_load(2021, week = 5)
plastics <- tuesdata$plastics
```

##### Data Wrangling
```{r}
#wrangling 2020 data
sum20_country <- plastics %>% 
  clean_names() %>% 
  filter(country != "EMPTY") %>%
  filter(year == "2020") %>% 
  mutate(grand_total = as.numeric(grand_total)) %>% 
  drop_na(grand_total) %>% 
  group_by(country) %>% 
  summarize(total_2020 = sum(grand_total))

#wrangling 2019 data
sum19_country <- plastics %>% 
  clean_names() %>% 
  filter(country != "EMPTY") %>%
  filter(year == "2019") %>% 
  mutate(grand_total = as.numeric(grand_total)) %>% 
  drop_na(grand_total) %>% 
  group_by(country) %>% 
  summarize(total_2019 = sum(grand_total))

#combining data frames and keeping top 10 countries with both years of data
combine_df <- right_join(sum19_country, sum20_country, by = "country") %>% 
  na.omit() %>% 
  mutate(difference = total_2020 - total_2019, # how to find the difference between to columns
         diff_group = as.factor(ifelse(total_2020 - total_2019 > 0, "Positive", "Negative"))) %>% #made a new column that stated whether the the grand total increased (positive) or decreased from 2019 to 2020
  top_n(abs(difference), n =10)

```

#### Data Visualization
```{r}
plot <- ggplot(data = combine_df, aes(y = reorder(country, abs(difference)), x = total_2019, xend = total_2020, 
                                      colour = diff_group)) +
  geom_dumbbell(size = 1.5,
                size_x = 3,
                size_xend = 3,
                colour_x = "cadetblue4",
                colour_xend = "blue") +
  scale_colour_manual(values = c("purple", "magenta")) +
  scale_x_continuous(breaks = c(seq(0, 60000, 10000))) +
  labs(title = "Change in Global Plastic Pollution",
       subtitle = "Top ten countries with the greatest change in plastic pollution between <span style = 'color:cadetblue4'><b>2019</b></span> and <span style = 'color:blue'><b>2020</b></span>. <span style = 'color:magenta'><b>Increase</b></span> in pollution was less common than <span style = 'color:purple'><b>decline</b></span>. <br>",
       caption = "\nSource: Break Free From Plastic | Visualization @nxrunning")+
  theme(plot.background = element_rect(fill = "aliceblue", colour = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.title = element_blank(),
        legend.position = "none",
        axis.ticks = element_line(colour = "antiquewhite4"),
        axis.text = element_text(colour = "antiquewhite4", size = 10, face = "bold"),
        plot.title = element_text(colour = "cornsilk4", size = 20, face = "bold"),
        plot.title.position = "plot",
        plot.subtitle = element_textbox_simple(
          size = 11,
          color = "antiquewhite4",
          padding = margin(5,5,5,5),
          margin = margin(0,0,0,0)),
        plot.caption = element_text(colour = "antiquewhite2", size = 10, face = "bold", hjust = 0))

plot
```

Experience with using a dumbbell plot to show change over time. This was the first time I used an online resource to find new skills and new techniques to visualize data. Tidy Tuesday is a great change to practive data visualization skills and learn new and interesting ways to present data.
